
This code is written as submission for the practical machine learning course provided by Johns Hopkins Univ via Coursera.
Below is the list of the packages that have been used.

```{r}
library(Hmisc)
library(foreach)
library(doParallel)
library(caret)
library(randomForest)
set.seed(2048)
options(warn=-1)
```
The data is loaded here.
The data was provided by coursera itself.


```{r}
training <- read.csv("C:/Users/prashant.mudgal/Documents/GitHub/practicalMachineLearning-ProgrammingAssignment/pml-training.csv", na.strings=c("#DIV/0!") )
nonTraining <- read.csv("C:/Users/prashant.mudgal/Documents/GitHub/practicalMachineLearning-ProgrammingAssignment/pml-testing.csv", na.strings=c("#DIV/0!") )

```
Column casting is done here

```{r}
for(i in c(8:ncol(training)-1)) {training[,i] = as.numeric(as.character(training[,i]))}

for(i in c(8:ncol(nonTraining)-1)) {nonTraining[,i] = as.numeric(as.character(nonTraining[,i]))}
```
Data filtering was done here and few columns were removed from dataset

```{r}
feature_set <- colnames(training[colSums(is.na(training)) == 0])[-(1:7)]
model_data <- training[feature_set]
feature_set
```
The model data from the filtered data

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

Multithreaded random forests. 5 RF with 150 trees.
Multithreading code was taken from internet
```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Error values and the confusion matrices for the training and the test datsets
```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```
Conclusions and Submission
--------------------------------

From confusion matrix values we can see that the model is quite accurate.

Submitting using COURSERA code

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- nonTraining
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```

